Counting infinitely by Oritatami co-transcriptional folding

### 2019-06-14 paper.tex
prelim copied..

%\bibliographystyle{splncs03}
%\bibliography{sofsem2020}

>>>>

%copy from TAMC2019 paper

Let $\Sigma$ be a finite set of types of abstract molecules, or \textit{beads}. 
A bead of type $a \in \Sigma$ is called an $a$-bead. 
By $\Sigma^*$ and $\Sigma^\omega$, we denote the set of finite sequences of beads and that of one-way infinite sequences of beads, respectively. 
The empty sequence is denoted by $\lambda$. 
Let $w = b_1 b_2 \cdots b_n \in \Sigma^*$ be a sequence of length $n$ for some integer $n$ and bead types $b_1, \ldots, b_n \in \Sigma$. 
The \textit{length} of $w$ is denoted by $|w|$, that is, $|w| = n$. %ここから要らない？↓やっぱり要る
For two indices $i, j$ with $1 \le i \le j \le n$, we let $w[i..j]$ refer to the subsequence $b_i b_{i+1} \cdots b_{j-1}b_j$; if $i = j$, then $w[i..i]$ is simplified as $w[i]$. 
For $k \ge 1$, $w[1..k]$ is called a \textit{prefix} of $w$.

Oritatami systems fold their transcript, which is a sequence of beads, over the triangular grid graph $\mathbb{T} = (V, E)$ cotranscriptionally. %↓
We designate one point in $V$ as the origin $O$ of $\mathbb{T}$. 
For a point $p \in V$, let $\hexagon_p^d$ denote the set of points which lie in the regular hexagon of radius $d$ centered at the point $p$. 
Note that $\hexagon_p^d$ consists of $3d(d+1)+1$ points. %←
A directed path $P = p_1 p_2 \cdots p_n$ in $\mathbb{T}$ is a sequence of \textit{pairwise-distinct} points $p_1, p_2, \ldots, p_n \in V$ such that $\{p_i, p_{i+1}\} \in E$ for all $1 \le i < n$. 
Its $i$-th point is referred to as $P[i]$. 
Now we are ready to abstract RNA single-stranded structures in the name of conformation. 
A \textit{conformation} $C$ (over $\Sigma$) is a triple $(P, w, H)$ of a directed path $P$ in $\mathbb{T}$, $w \in \Sigma^*$ of the same length as $P$, and a set of h-interactions $H \subseteq \bigl\{\{i, j\} \bigm| 1 \le i, i+2 \le j, \{P[i], P[j]\} \in E \bigr\}$. 
This is to be interpreted as the sequence $w$ being folded along the path $P$ in such a manner that its $i$-th bead $w[i]$ is placed at the $i$-th point $P[i]$ and the $i$-th and $j$-th beads are bound (by a hydrogen-bond-based interaction) if and only if $\{i, j\} \in H$. 
The condition $i+2 \le j$ represents the topological restriction that two consecutive beads along the path cannot be bound. 
The \textit{length} of $C$ is defined to be the length of its transcript $w$ (that is, equal to the length of the path $P$). 
A \textit{rule set} $R \subseteq \Sigma \times \Sigma$ is a symmetric relation over $\Sigma$, that is, for all bead types $a, b \in \Sigma$, $(a, b) \in R$ implies $(b, a) \in R$.
A bond $\{i, j\} \in H$ is \textit{valid with respect to $R$}, or simply $R$-valid, if $(w[i], w[j]) \in R$. 
This conformation $C$ is \textit{$R$-valid} if all of its bonds are $R$-valid. %↓
For an integer $\alpha \ge 1$, $C$ is \textit{of arity $\alpha$} if it contains a bead that forms $\alpha$ bonds but none of its beads forms more. 
By $\mathcal{C}_{\le \alpha}(\Sigma)$, we denote the set of all conformations over $\Sigma$ whose arity is at most $\alpha$; its argument $\Sigma$ is omitted whenever $\Sigma$ is clear from the context.

The oritatami system grows conformations by an operation called elongation. 
Given a rule set $R$ and an $R$-valid conformation $C_1 = (P, w, H)$, we say that another conformation $C_2$ is an elongation of $C_1$ by a bead $b \in \Sigma$, written as $C_1 \xrightarrow{R}_b C_2$, if $C_2 = (P p, wb, H \cup H')$ for some point $p \in V$ not along the path $P$ and set $H' \subseteq \bigl\{ \{i, |w|+1\} \bigm| 1 \le i < |w|, \{P[i], p\} \in E, (w[i], b) \in R \bigr\}$ of bonds formed by the $b$-bead; this set $H'$ can be empty. 
Note that $C_2$ is also $R$-valid. 
This operation is recursively extended to the elongation by a finite sequence of beads as: for any conformation $C$, $C \xrightarrow{R}_\lambda^* C$; and for a finite sequence of beads $w \in \Sigma^*$ and a bead $b \in \Sigma$, a conformation $C_1$ is elongated to a conformation $C_2$ by $wb$, written as $C_1 \xrightarrow{R}_{wb}^* C_2$, if there is a conformation $C'$ that satisfies $C_1 \xrightarrow{R}_w^* C'$ and $C' \xrightarrow{R}_b C_2$.

An \textit{oritatami system} (OS) $\Xi = (\Sigma, R, \delta, \alpha, \sigma, w)$ is composed of
\begin{itemize}
\item a set $\Sigma$ of bead types, 
\item a rule set $R \subseteq \Sigma \times \Sigma$, 
\item a positive integer $\delta$ called the \textit{delay}, 
\item a positive integer $\alpha$ called the \textit{arity}, %←arityの行だけ要らない？
\item an initial $R$-valid conformation $\sigma \in C_{\le \alpha}(\Sigma)$ called the \textit{seed}, whose first bead is assumed to be at the origin $O$ without loss of generality, 
\item a (possibly infinite) \textit{transcript} $w \in \Sigma^* \cup \Sigma^\omega$, which is to be folded upon the seed by stabilizing beads of $w$ one at a time so as to minimize energy collaboratively with the succeeding $\delta{-}1$ nascent beads. 
\end{itemize}
%
The energy of a conformation $C = (P, w, H)$, denoted by $\Delta G(C)$, is defined to be ${-}|H|$; the more bonds a conformation has, the more stable it gets. 
The set $\mathcal{F}(\Xi)$ of conformations \textit{foldable} by the system $\Xi$ is recursively defined as: the seed $\sigma$ is in $\mathcal{F}(\Xi)$; and provided that an elongation $C_i$ of $\sigma$ by the prefix $w[1..i]$ be foldable (i.e., $C_0 = \sigma$), its further elongation $C_{i+1}$ by the next bead $w[i+1]$ is foldable if 
\begin{equation}\label{eq:OS_CF}
C_{i+1} \in \argmin_{
\substack{
C \in \mathcal{C}_{\le \alpha} s.t. \\
C_i \xrightarrow{R}_{w[i+1]}C \\
}
}
\min \Bigl\{ \Delta G(C') \Bigm|
C \xrightarrow{R}^*_{w[i+2...i+k]}C', k\le \delta, C' \in \mathcal{C}_{\le \alpha}
\Bigr\}.
\end{equation}
%
Then we say that the bead $w[i+1]$ and the bonds it forms are \textit{stabilized} according to $C_{i+1}$. 
The easiest way to understand this stabilization process should be the video available at \href{https://www.dailymotion.com/video/x3cdj35}{\tt https://www.dailymotion.com/video/x3cdj35}, in which the Turing universal oritatami system by Geary et al. \cite{GeMeScSe2018}, whose delay is 3, is running. 
Note that an arity-$\alpha$ oritatami system cannot fold any conformation of arity larger than $\alpha$. %←
A conformation foldable by $\Xi$ is \textit{terminal} if none of its elongations is foldable by $\Xi$. 
%The oritatami system $\Xi$ is \textit{deterministic} if for all $i \ge 0$, there exists at most one $C_{i+1}$ that satisfies \eqref{eq:OS_CF}. 
%A deterministic oritatami system folds into a unique terminal conformation. 
%An oritatami system with the empty rule set just folds into an arbitrary elongation of its seed nondeterministically. 
%Thus, the rule set is reasonably assumed non-empty.

\bibliographystyle{splncs03}
\bibliography{sofsem2020}
